Proposal: A Natural Language Programming Toolkit for Python

### 2001-03-03 technical/proposal/proposal.tex
Moved the task-oriented module prototype table.  Minor wording change.
svn/trunk@11
Added "content" and "nomenclature" subsubsections to the module design subsection.  Gave URL for reference docs.  Minor doc changes.  Added docs for FreqDistI and ProbDistI.  Added docs for TaggerI.  Added bibliography items for some NLP books.
svn/trunk@13

@@

Epydoc is currently under development.  Contact Edward Loper for more
information about Epydoc.  For examples of the HTML pages generated by 
Epydoc, see the reference documentation for the toolkit's prototype
modules:

 \texttt{http://www.cis.upenn.edu/$\sim$edloper/??}

@@

As part of the preliminary design process, I desinged and implemented
six prototype modules.  These modules are summarized in figures
\ref{fig:mod-data} and \ref{fig:mod-task}, and described in the
following sections.  Most of these summaries are copied directly from
the modules' reference documentation.  For more information, see the
reference documentation at:

 \texttt{http://www.cis.upenn.edu/$\sim$edloper/??}

@@

\begin{figure}
\noindent
\begin{centering}
\begin{boxedminipage}{.95\textwidth}
\begin{itemize}
  \item \textbf{Tokenizers}: Separate a string of text into a list of
       Tokens. 
  \begin{itemize}
     \item \textit{TokenizerI}
  \end{itemize}

  \item \textbf{Taggers}: Assign tags to each Token in a list of Tokens.
  \begin{itemize}
     \item \textit{TaggerI}
  \end{itemize}
\end{itemize}
\end{boxedminipage}\\
\end{centering}
 \caption[Task-oriented module prototypes]{Task-oriented module
 prototypes.  Under each module, the interfaces and top-level classes
 defined by that module are listed.}
\label{fig:mod-task}
\end{figure}

@@

    An unordered container class that contains no duplicate elements.
    In particular, a set contains no elements $e_1$ and $e_2$ such
    that $e_1=e_2$.  Currently, the \texttt{Set} class is given a
    fairly minimal implementation.  However, more members may be
    defined in the future.

@@

A Token is an occurance of a single unit of text, such as a word or a
punctuation mark.  The token module defines the \texttt{TokenTypeI}
interface, the \texttt{TextLocationI} interface, and the \texttt{Token
class}.

@@

    Token types can be compared for equality, and should be equal if
    and only if they refer to the same token type.

@@

\subsubsection{\texttt{Token} Class}

    A \texttt{Token} is an occurance of a single unit of text, such as
    a word or a punctuation mark.  A Token consists of a token type
    and a source.  The token type is the unit of text (e.g., a
    specific word).  The source is the position at which this token
    occured in the text.

@@

@@

A syntax node is a single internal node of a syntax tree.  Each syntax
node is characterized by a tag and an ordered list of children.  The
tag specifies the type of the syntax Node (e.g., ``NP'' or ``V'').
Each child must either be a \texttt{Token}, or implement the
\texttt{SyntaxNodeI} interface.  Access to the children is implemented
by emulating a sequence type.  The \texttt{SyntaxNodeI} does
\emph{not} perform any semantic or syntactic checks on its arguments.
Figure \ref{fig:syntaxnode} shows how syntax nodes can be used:

@@

    An event is a subset of some sample space.  Note that this subset
    need not be finite.  Events are typically written as the set of
    samples they contain, or as a function in first order logic.
    Examples are:

@@

    Classes implementing the \texttt{EventI} interface may choose
    to only support certain classes of samples, or may choose to only
    support certain types of events as arguments to the optional
    methods (\texttt{\_\_cmp\_\_}, \texttt{subset}, etc.).  If a
    method is unable to return a correct result because it is given an 
    unsupported type of sample or event, it should raise a
    NotImplementedError.

@@

(still writing documentation)

@@

(still implementing)

@@

The process of separating a string of text into Tokens is called
\emph{tokenization}.  The tokenizer module defines the
\texttt{TokenizerI} interface, which should be implemented by classes
that tokenize text strings.

@@

    A tokenizer is a processing class responsible for separating a
    string of text into a list of \texttt{Token}s.  This process
    is also known as \emph{tokenizing} the string of text.  Particular
    \texttt{Tokenizer}s may split the text at different points, or
    may produce Tokens with different types of \texttt{TokenType}.

@@

The process of assigning a tag to each Token in a

@@

@@

@@

>>>>

\subsubsection{Content}

The decision of what modules to include in the toolkit will be
initially guided by the past content of CIS-530 \cite{cis530}.  Later,
the content of several Natural Language Processing text books may be
used to guide decisions about what modules to write \cite{allen1995},
\cite{gazdar1989}, \cite{jurafsky2000}, \cite{manning2000}.

\subsubsection{Nomenclature}

When learning to use the toolkit, students intuitions about each
entity in the toolkit will be guided by the entities names'.  It is
therefore important to use consistant, correct, and intuitive
nomenclature when naming the entities defined by the toolkit.
The documentation should be careful to always use correct
nomenclature; in particular, it should never conflate two concepts
with the same name.

@@

Epydoc is currently under development.  Contact Edward Loper for more
information about Epydoc.  For examples of the HTML pages generated by
Epydoc, see the reference documentation for the toolkit's prototype
modules 
at ``\texttt{http://www.cis.upenn.edu/$\sim$edloper/nltk/nltk.html}''.

@@

As part of the preliminary design process, I desinged and implemented
six prototype modules.  These modules are summarized in figures
\ref{fig:mod-data} and \ref{fig:mod-task}, and described in the
following sections.  Most of these descriptions are copied directly from
the modules' reference documentation.  For more information, see the
reference documentation 
at ``\texttt{http://www.cis.upenn.edu/$\sim$edloper/nltk/nltk.html}''.

@@

@@

    A \texttt{Set} is an unordered container class that contains no
    duplicate elements.  In particular, a set contains no elements
    $e_1$ and $e_2$ such that $e_1=e_2$.  Currently, the \texttt{Set}
    class is given a fairly minimal implementation.  However, more
    members may be defined in the future.

@@

A \texttt{Token} is an occurance of a single unit of text, such as a word or a
punctuation mark.  The token module defines the \texttt{TokenTypeI}
interface, the \texttt{TextLocationI} interface, and the \texttt{Token}
class.

@@

    \texttt{TokenType}s can be compared for equality, and should be
    equal if and only if they refer to the same token type.

@@

\subsubsection{Token Class}

    A \texttt{Token} is an occurance of a single unit of text, such as
    a word or a punctuation mark.  A \texttt{Token} consists of a token type
    and a source.  The token type is the unit of text (e.g., a
    specific word).  The source is the position at which this token
    occured in the text.

@@

This module is still under development.  Several open questions
remain, such as whether \texttt{SyntaxTree}s should correspond to
individual occurances of syntax trees, or to ``types'' of syntax trees
(c.f. the distinction between \texttt{Token} and TokenType).

@@

A syntax node is a single internal node of a syntax tree.  Each syntax
node is characterized by a tag and an ordered list of children.  The
tag specifies the type of the syntax Node (e.g., ``NP'' or ``V'').
Each child must either be a \texttt{Token}, or implement the
\texttt{SyntaxNodeI} interface.  Access to the children is implemented
by emulating a sequence type.  The \texttt{SyntaxNodeI} does
\emph{not} perform any semantic or syntactic checks on its arguments.
Figure \ref{fig:syntaxnode} shows how syntax nodes can be used.

@@

    An \texttt{Event} is a subset of some sample space.  Note that
    this subset need not be finite.  Events are typically written as
    the set of samples they contain, or using a function.  Examples
    are:

@@

    Classes implementing the \texttt{EventI} interface may choose
    to only support certain classes of samples, or may choose to only
    support certain types of events as arguments to the optional
    methods (\texttt{\_\_cmp\_\_}, \texttt{subset}, etc.).  If a
    method is unable to return a correct result because it is given an 
    unsupported type of sample or event, it should raise a
    \texttt{NotImplementedError}.

@@

    A \texttt{FreqDist} represents the frequency distribution for the
    outcomes of an experiment.  A frequency distribution records the
    number of times each outcome of an experiment has occured.  For
    example, a frequency distribution could be used to record the
    frequency of each token in a document.  Formally, a frequency
    distribution can be defined as a function mapping from samples to
    the number of times that sample occured as an outcome.

    Frequency distributions are generally constructed by running a
    number of experiments, and incrementing the count for a sample
    every time it is an outcome of an experiment.  For example, the
    following code will produce a frequency distribution that encodes
    how often each word occurs in a text:

\begin{verbatim}
freqDist = SimpleFreqDist()
for word in document:
    freqDist.inc(word)
\end{verbatim}

    Classes implementing the \texttt{FreqDistI} interface may
    choose to only support certain classes of samples or events.  If a
    method is unable to return a correct result because it is given an
    unsupported type of sample or event, it should raise a
    \texttt{NotImplementedError}.

    Since several methods defined by \texttt{FreqDistI} can accept
    either events or samples, classes that implement the EventI
    interface should never be used as samples for a frequency
    distribution.

    Frequency distributions are required to implement the methods
    \texttt{inc()}, \texttt{N()}, \texttt{count()},
    \texttt{freq()}, \texttt{max()}, \texttt{cond\_freq()}, 
    and \texttt{cond\_max()}.  In the future, this list may be
    exapanded, and optional methods may be added.

\vspace{2mm}\noindent
\textit{The \texttt{FreqDistI} interface is currently implemented by the
classes \texttt{SimpleFreqDistI} and \texttt{CFFreqDistI}.}

@@

    A probability distribution for the outcomes of an experiment.  A
    probability distribution specifies how likely it is that an
    experiment will have any given outcome.  For example, a
    probability distribution could be used to predict the probability
    that a given word will appear in a given context.  Formally, a
    probability distribution can be defined as a function mapping from
    samples to nonnegative real numbers, such that the sum of every
    number in the function's range is 1.0.  \texttt{ProbDist}s are
    often used to model the probability distribution underlying a
    frequency distribution.

    Classes implementing the \texttt{ProbDistI} interface may
    choose to only support certain classes of samples or events.  If a
    method is unable to return a correct result because it is given an
    unsupported type of sample or event, it should raise a
    NotImplementedError.

    Since several methods defined by \texttt{ProbDistI} can accept
    either events or samples, classes that implement the EventI
    interface should never be used as samples for a probability
    distribution.

    Probability distributions are required to implement the methods
    \texttt{prob()} and \texttt{cond\_prob()},
    \texttt{max()}, and \texttt{cond\_max()}.  In the future,
    this list may be exapanded, and optional methods may be added.

\vspace{2mm}\noindent
\textit{The \texttt{ProbDistI} interface is not currently implemented
any class.}

\begin{figure}
\noindent
\begin{centering}
\begin{boxedminipage}{.95\textwidth}
\begin{itemize}
  \item \textbf{Tokenizers}: Separate a string of text into a list of
       \texttt{Token}s. 
  \begin{itemize}
     \item \textit{TokenizerI}
  \end{itemize}

  \item \textbf{Taggers}: Assign tags to each \texttt{Token} in a list of \texttt{Token}s.
  \begin{itemize}
     \item \textit{TaggerI}
  \end{itemize}
\end{itemize}
\end{boxedminipage}\\
\end{centering}
 \caption[Task-oriented module prototypes]{Task-oriented module
 prototypes.  Under each module, the interfaces and top-level classes
 defined by that module are listed.}
\label{fig:mod-task}
\end{figure}

@@

The process of separating a string of text into \texttt{Token}s is called
\emph{tokenization}.  The tokenizer module defines the
\texttt{TokenizerI} interface, which should be implemented by classes
that tokenize text strings.

@@

    A \texttt{Tokenizer} is a processing class responsible for
    separating a string of text into a list of \texttt{Token}s.  This
    process is also known as \emph{tokenizing} the string of text.
    Particular \texttt{Tokenizer}s may split the text at different
    points, or may produce \texttt{Token}s with different types of
    \texttt{TokenType}.

@@

The process of assigning a descriptive tag to each \texttt{Token} in a
list of \texttt{Token}s is called \emph{tagging}.  Often, tagging is
used to assign a part-of-speech tag to each \texttt{Token} in a list
of \texttt{Token}s.  Particular \texttt{Taggers} may attatch different
types of tags to the \texttt{Token}s in the list, or may use different
algorithms for deciding which tag to attatch.

@@

    A \texttt{Tagger} is a processing class for assigning a tag to
    each token in an ordered list of tokens.  Taggers are required to
    define one function, \texttt{tag}, which takes a list of
    \texttt{Token}s, and returns a list of \texttt{Token}s.
    Typically, the input tokens will be \texttt{SimpleToken}s, and the
    output tokens will be \texttt{TaggedToken}s.  However, taggers may
    also be written to map between other types of tokens, as long as
    they are still performing the same conceptual task.

    Classes implementing the \texttt{TaggerI} interface may choose to
    only support certain classes of tokens.  If a method is unable to
    return a correct result because it is given an unsupported class
    of token, then it should raise a \texttt{NotImplementedError}.

@@

\bibitem{allen1995}
{Allen, James.  1995.  
 \emph{Natural Language Understanding}.
 Redwood, CA: Benjamin Cummings.}

@@

\bibitem{gazdar1989}
{Gazdar, Gerald and Mellish, Chris.  1989.
 \emph{Natural Language Processing in LISP : An Introduction to
       Computational Linguistics}. 
 Reading, MA: Addison-Wesley Pub. Co.}

\bibitem{jurafsky2000}
{Jurafsky, Daniel and Martin, James H.  2000.
 \emph{Speech and Language Processing: An Introduction to Natural
       Language Processing, Computational Linguistics, and Speech
       Recognition}.
 Upper Saddle River, NJ: Prentice Hall.}

\bibitem{manning2000}
{Manning, Christopher D. and Schutze, Hinrich.  2000.  
 \emph{Foundations of Statistical Natural Language Processing}.
 Cambridge, MA: MIT Press.}
